---
title: "covid 19"
author: "Emraida Ortanez"
date: "10/19/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  

```{r}
barchartGC(~result + Sex,data=kRdata,type="percent" ,main= "Frequency Bar Chart")
```
```{r}
barchartGC(~result + Sex,data=kRdata, main = "Frequency Bar Chart")
```
   
 According the frequency bar chart there are more women alive than the men and the second bar shows that more men are dead than women
The graph also shows that about 62 to 65 percent of men are dead and about 35 to 38 percent of women are dead.
  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
The result shows that about 64 men out of 100 have died of covid 19 compared to the result of 36 out of 100 women who have died . 
his gives a ratio of about 1:3 when simplified .    

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



## Conclusions

